# (PART) Praxis {-} 

# Use of flow rate equations for sizing valves

## Example 1. 

Sizing a valve without fittings, constant flow, constant upstream, and constant downstream pressures.

   - Step 1. Specify the variables required  to size the valve.
   - Step 2. Read the values of the selected valve.
   - Step 3. Determine the flow coefficient $k_v$, the opening degree for the flow and $F_L$.
   - Step 4. calculate the value of sigma ($\sigma$) and its boundaries.
  
### Step 1. Specify the variables required  to size the valve.

```{r include=FALSE}

  # Basic information
  temp <-  23.300   # 째C
  elev <-   0.000   # m 
  p1   <-   5.700   # bar
  p2   <-   4.800   # bar
  flow <- 795.600   # m3/h
  d    <-   0.250   # m (DN of the valve)
  D1   <-   0.250   # m (DN Pipe Upstream)
  D2   <-   0.250   # m (DN Pipe Downstream)
  
  # Control Characteristics
  cylinder	<- "Equal Percentage"
  kv.b      <- -3.893
  kv.d	    <-  1.146
  kv.e      <- 61.038
  zvs	      <-  3.000
  fls	      <-  0.598
  
  # Calculation
  kvs       <- kv_value(d, zvs)
  kv_needed <- kv(p1, p2, flow, temp)
  kv_kvs    <- kv_needed/kvs*100
  position  <- inv_LL3(kv_needed/kvs, kv.b, kv.d, kv.e)
  fl        <- fl_function(position, kv.b, kv.d, kv.e, fls)
  zeta      <- zeta_vaule(d, kv_needed)
  sig_1     <- sigma_1(p1, p2, elev, temp)

  Sig_i     <- Sigma_i(position,  kv.b,  kv.d,  kv.e, fls) # Incipient Cavitation
  Sig_c     <- Sigma_c(position,  kv.b,  kv.d,  kv.e, fls) # Constant Cavitation
  Sig_mv    <- Sigma_mv(position,  kv.b,  kv.d,  kv.e, fls) #  Maximum Vibration Cavitation


  # Table for the plot
  df <- tibble( factor = c( "Kv", "Kv_Kvs", "Fl", "zeta",
                            "sig_1", "Sig_i", "Sig_c", "Sig_mv"), 
                     y = c( kv_needed, kv_kvs, fl, zeta, 
                            sig_1, Sig_i, Sig_c, Sig_mv ), 
                     x = c( position, position, position, position,
                            position, position, position, position))
  
```

Service data : 
  1. Water pressure must be managed from 30 bars to 6 bars.
  2. The control valve has a diameter of `r round(d*1000,0)` mm and is smaller than the diameters of the pipe upstream (`r round(D1*1000,0)` mm) and downstream (`r round(D2*1000,0)` mm)  of the valve.


physical constants:
  
  - Fluid : Water 
  - Water Temperature :   `r temp` 째C
  - Altitude:             `r elev` msal.
  - Atmosphere Pressure : `r atm_pressure(masl = elev)` bar
  - Vapor Pressure :      `r vapour_pressure(temp)*0.01` bar
  - Kinematic viscosity : `r scientific(kinematic_viscosity(temp))` $(mm^2/s)$

Service Conditions:

  - $P_1 =$                         `r p1` bar
  - $P_2 =$                         `r p2` bar
  - $\Delta{P}$                     `r p1 - p2` bar
  - $q =$                           `r (flow)` $m^3/h$ 
  - Diameter Pipe Upstream $D_1$:   `r round(D1*1000,0)` mm  
  - Diameter Pipe Downstream $D_2$: `r round(D2*1000,0)` mm  

Valve Design (first approx.):
  
  - Valve: Plunger Valve
  - Diameter Valve $d$: `r round(d*1000,0)` mm  
  - PN-10
  - Control Element : Cylinder type `r cylinder`
  - Control Characteristic: 


### Step 2. Read the values of the selected Control Characteristic.

check the values of the preselected valve (values supplied by the valve producer). 
  
  - With the valve 100% open: 
    - Zeta Value $\zeta_{vs}$ :                       `r comma(zvs)`
    - Flow coefficient : $K_{vs}$ :                   `r comma(kvs)` $m^3/h$
    - liquid pressure recovery factor ${F_{L}}_{s}$ : `r fls`


### Step 3. Calculate the required flow coefficient ($K_v$).

Using the function \@ref(eq:Kv-01):

$$
  K_v = q \cdot \sqrt{\frac {(\rho / \rho_{0})}{\Delta P}}
$$
the Required $K_v$ is `r comma(kv_needed)` $m^3/h$ and the relative flow coefficient is $K_v/K_{vs}=$ `r percent(kv_kvs/100)`.

Entering the graph of $k_V/K_{VS}$ with the value of `r percent(kv_kvs/100)` we find the position of the valve required for the flow of `r comma(flow)` $m^3/h$ and $\Delta{P}$ of `r p1 - p2` bar is `r percent(position/100)`.


```{r kv-kvs-Example-1, echo=FALSE, fig.cap="Inherent Valve Characteristics"}
  
  data_points <- df %>%  filter(factor == "Kv_Kvs")
  
  arrows_points <- tibble( x0 = 0,        y0 = kv_kvs,
                           x1 = position, y1 = kv_kvs,
                           x2 = position, y2 = 0)
  

  plot_kv_kvs(kv.b, kv.d, kv.e, cylinder) +
    geom_point( data = data_points, aes( y = y, x = x ), 
                colour = "red", size = 3) +
     geom_segment(data = arrows_points, 
                  aes(x = x0, y = y0, xend = x1, yend = y1),
                  arrow = arrow(length = unit(0.03, "npc"), type = "closed"))+
     geom_segment(data = arrows_points, 
                  aes(x = x1, y = y1, xend = x2, yend = y2),
                  arrow = arrow(length = unit(0.03, "npc"), type = "closed"))
  
```

Or it can be done directly with the $K_V =$ `r (kv_needed)` $m^3/h$  value of on the $K_V$ graph.

```{r kv-Example-1, echo=FALSE, fig.cap="Flow coefficient"}
  
  data_points <- df %>%  filter(factor == "Kv")
  
  arrows_points <- tibble( x0 = 0,        y0 = kv_needed,
                           x1 = position, y1 = kv_needed,
                           x2 = position, y2 = 0)


  plot_kv(kv.b, kv.d, kv.e, d, zvs, cylinder) +
    geom_point( data = data_points, aes( y = y, x = x ), 
                colour = "red", size = 3) +
     geom_segment(data = arrows_points, 
                  aes(x = x0, y = y0, xend = x1, yend = y1),
                  arrow = arrow(length = unit(0.03, "npc"), type = "closed"))+
     geom_segment(data = arrows_points, 
                  aes(x = x1, y = y1, xend = x2, yend = y2),
                  arrow = arrow(length = unit(0.03, "npc"), type = "closed"))

```


Knowing that the valve position is `r round(position,2)`%, it is possible to read the Zeta Value $\zeta_{v}$ and the Liquid Pressure Recovery Factor ${F_L}$.:

  - $\zeta_{v}=$ `r comma(zeta) `
  - ${F_L}=$ `r round(fl,3) `
  - Reynolds Number of the valve `r comma(Reynolds_valve(flow, kv_needed, d*1000, D1*1000, D2*1000, temp, fl, 1)) `


```{r zv-Example-1, echo=FALSE, fig.cap="Zeta Value"}

  data_points <- df %>%  filter(factor == "zeta")


  arrows_points <- tibble( x0 = position, y0 = 0,
                           x1 = position, y1 = zeta,
                           x2 = 0,        y2 = zeta)


  plot_zv(kv.b, kv.d, kv.e, zvs, cylinder) +
    geom_point( data = data_points, aes( y = y, x = x ), 
                colour = "red", size = 3)  +
     geom_segment(data = arrows_points, 
                  aes(x = x0, y = y0, xend = x1, yend = y1),
                  arrow = arrow(length = unit(0.03, "npc"), type = "closed"))+
     geom_segment(data = arrows_points, 
                  aes(x = x1, y = y1, xend = x2, yend = y2),
                  arrow = arrow(length = unit(0.03, "npc"), type = "closed"))
```


```{r fl-Example-1, echo=FALSE, fig.cap="Liquid pressure Recovery Factor"}

  data_points <- df %>%  filter(factor == "Fl")

  arrows_points <- tibble( x0 = position, y0 = 0,
                           x1 = position, y1 = fl,
                           x2 = 0,        y2 = fl)

  plot_fl( kv.b, kv.d, kv.e, fls, cylinder) +
    geom_point( data = data_points, aes( y = y, x = x ), 
                colour = "red", size = 3) +
     geom_segment(data = arrows_points, 
                  aes(x = x0, y = y0, xend = x1, yend = y1),
                  arrow = arrow(length = unit(0.03, "npc"), type = "closed"))+
     geom_segment(data = arrows_points, 
                  aes(x = x1, y = y1, xend = x2, yend = y2),
                  arrow = arrow(length = unit(0.03, "npc"), type = "closed"))
```

### Step 4. calculate the value of sigma ($\sigma$) and its boundaries.

Using the function \@ref(eq:sigma-value-01):

$$
  \sigma_{1} = \frac{P_2-P_v}{\Delta P_{net}}
$$

  - Sigma value ($\sigma_{1}$): `r round(sig_1,3)`

  - Incipient Cavitation ($\sigma_{i}$): `r round(Sig_i,3)`
  
  - Constant Cavitation ($\sigma_{c}$):  `r round(Sig_c,3)`

  - Maximum Vibration Cavitation ($\sigma_{mv}$): `r round(Sig_mv,3)`



```{r sigma-Example-1, echo=FALSE, warning=FALSE, fig.cap="Sigma"}
  
  # https://sebastiansauer.github.io/dplyr_filter/
  
  data_points <- df %>%  
    filter(factor %in% c("sig_1", "Sig_i", "Sig_c", "Sig_mv"))

  arrows_points <- tibble( x0 = 0,         y0 = sig_1,
                           x1 = position,  y1 = sig_1,
                           x2 = 0,         y2 = Sig_i,
                           x3 = position,  y3 = Sig_i,
                           x4 = 0,         y4 = Sig_c,
                           x5 = position,  y5 = Sig_c,
                           x6 = 0,         y6 = Sig_mv,
                           x7 = position,  y7 = Sig_mv,
                           x8 = position,  y8 = 0)

  # Line type : "solid", "dashed", "dotted", "dotdash", "longdash", "twodash")
  
  lty <- "dotted"
  
  plot_sigma( kv.b, kv.d, kv.e, fls, cylinder) +
    geom_point( data = data_points, aes( y = y, x = x ), 
                colour = "black", size = 2) +
     geom_segment(data = arrows_points, aes(x = x0, y = y0, xend = x1, yend = y1), linetype = lty) +
     geom_segment(data = arrows_points, aes(x = x1, y = y1, xend = x8, yend = y8), linetype = lty) +
     geom_segment(data = arrows_points, aes(x = x2, y = y2, xend = x3, yend = y3), linetype = lty) +
     geom_segment(data = arrows_points, aes(x = x4, y = y4, xend = x5, yend = y5), linetype = lty) +
     geom_segment(data = arrows_points, aes(x = x6, y = y6, xend = x7, yend = y7), linetype = lty)
    
  
```

## Example 2. 

   - Step 1. Specify the variables required  to size the valve.
   - Step 2. Determine the equation constants.
   - Step 3. Determine the piping geometry factor ($F_P$).
   - Step 4. Determine the pressure drop to use for sizing ($\Delta{P_{sizing}}).$
   - Step 5. Calculate the required flow coefficient ($K_v$).

### Step 1. Specify the variables required  to size the valve.

```{r include=FALSE}

  library(tidyverse)
  library(wcontrolvalve)

  temp <- 20.00   # 째C
  elev <- 500.0   # m 
  p1   <- 30.00   # bar
  p2   <- 6.000   # bar
  flow <- 0.250   # m3/s
  d    <- 0.300   # m
  D1   <- 0.500   # m (DN Pipe Upstream)
  D2   <- 0.400   # m (DN Pipe Downstream)
  
  # kv curve SZ-40
  
  kv.b	<- -2.755
  kv.d	<-  1.979
  kv.e	<- 99.215
  zvs	  <- 15.830
  fl	  <-  0.775
```

  1. Water pressure must be managed from 30 bars to 6 bars.
  2. The control valve has a diameter of `r round(d*1000,0)` mm and is smaller than the diameters of the pipe upstream (`r round(D1*1000,0)` mm) and downstream (`r round(D2*1000,0)` mm)  of the valve.

Service Conditions:

  - Fluid : Water 
  - Water Temperature : `r temp` 째C
  - Altitude: `r elev` msal.
  - $P_1 =$ `r p1` bar
  - $P_2 =$ `r p2` bar
  - $q =$ `r flow` LPS
  - Diameter Pipe Upstream $D_1$: `r round(D1*1000,0)` mm  
  - Diameter Pipe Downstream $D_2$: `r round(D2*1000,0)` mm  

Valve Design (first approx.):
  
  - Valve: Plunger Valve
  - Diameter Valve $d$: `r round(d*1000,0)` mm  
  - PN-40
  - Control Element : Cylinder SZ-40
  - Control Characteristic: Equal-Percentage Characteristic

### Step 2. Determine the equation constants.

  - 100% open 
    - Zeta Value $\zeta_{vs}$: `r zvs`
    - Flow coefficient: $K_{vs}$ : `r round(kv_value(d, zvs),0)` $m^3 / h$
    - liquid pressure recovery factor ${F_{L}}_{s}$ : `r fl`
  - Atmosphere Pressure : `r atm_pressure(masl = elev)` bar
  - Vapor Pressure : `r vapour_pressure(temp)*0.01` bar
  - Kinematic viscosity : `r kinematic_viscosity(temp)` $(m^2/s)*1e-6$

### Step 3. Determine the piping geometry factor ($F_P$).

```{r valve-with-fittings2, echo=FALSE, fig.cap = 'Valve with fittings [@wagner2008]', fig.width = 10, fig.asp = 0.7, fig.align = "center"}
  file_image <- here::here("image", "valve_with_fittings.PNG")
  knitr::include_graphics(file_image)
```

according to equation \@ref(eq:Fp) :

$$
  F_{P} = \frac{1}{\sqrt{1+\frac{\sum{\zeta} \cdot \left(\frac{K_v}{d^2} \right)^2}{0.0016}}}
$$
Where \@ref(eq:sumZ)

$$
  \sum{\zeta} = (\zeta_{1} + \zeta_{2}) + (\zeta_{B_1} - \zeta_{B_2})
$$

and

Resistance coefficient of the valve in the inlet \@ref(eq:Z1): 
    $\zeta_{1} = 0.5 \cdot \left( 1 - \left( \frac{d}{D_1}\right)^2 \right)^2 =$ `r round(0.5*(1-(d/D1)^2)^2,4)`
  
Resistance coefficient of the valve in the outlet \@ref(eq:Z2): 
    $\zeta_{2} = \left( 1 - \left( \frac{d}{D_2}\right)^2 \right)^2 =$ `r round((1-(d/D2)^2)^2,4)`
  
Bernoulli pressure number in the valve inlet \@ref(eq:Zb1): 
    $\zeta_{B_1} = 1- \left( \frac{d}{D_1}  \right)^4 =$  `r round((1-(d/D1))^4,4)`
  
Bernoulli pressure number in the valve outlet \@ref(eq:Zb2): 
    $\zeta_{B_2} = 1- \left( \frac{d}{D_2}  \right)^4 =$ `r round((1-(d/D2))^4,4)`

$\sum{\zeta} =$ `r round(0.5*(1-(d/D1)^2)^2 + (1-(d/D2)^2)^2 + (1-(d/D1))^4 - (1-(d/D2))^4, 4)`

$F_P =$ `r fp(kv_value(d, zvs), d, D1, D2)`

