# Example 2. 

The objective of this exercise is to select from three possible types of control valves (the same valve but different cylinders) the most suitable for the operation of the system.

the conditions are as follows:

  - The diameter of the valve is smaller than the diameter of the pipes, and the piping system requires a constant flow,
  - the pressures vary both upstream and downstream of the valve, and
  - Initially, only three operational points will be evaluated.
  
The steps to follow are those:

   - Step 1. Specify the variables required  to size the valve.
   - Step 2. Determine the equation constants.
   - Step 3. Determine the piping geometry factor ($F_P$).
   - Step 4. Determine the pressure drop to use for sizing ($\Delta{P_{sizing}}).$
   - Step 5. Calculate the required flow coefficient ($K_v$).

## Step 1. Specify the variables required  to size the valve.

```{r include=FALSE}

  temp <- 15.00   # °C
  elev <- 2700    # m 
  flow <- 9396    # m3/h
  d    <- 1.000   # m
  d1   <- 1.4     # m (DN Pipe Upstream)
  d2   <- 1.2     # m (DN Pipe Downstream)
  
  # base_data Pressures (bar)
  base_data <- tribble(
    ~measurement, ~p1,   ~p2, ~flow,
    "Min.",      1.720, 0.614,  9396,
    "Mean",      1.410, 0.684,  9396,
    "Max.",      1.370, 1.009,  9396
  ) 
  
  # Cylinder parameter/Characteristics
  
  cylinder <- tribble(
    ~typ,     ~kv.b,  ~kv.d,   ~kv.e,   ~zvs,  ~fls,
    "typ_01", -2.393,	1.389,  67.422,	 1.780,	0.617, # 0
    "typ_02", -3.893,	1.146,  61.038,  3.000,	0.598, # 20-30
    "typ_03", -2.755, 1.979,  99.215, 15.830, 0.775  # 40
  ) 
  
  cylinder <- cylinder %>% 
    mutate( kvs  = kv_value(d, zvs)) %>% 
    mutate( fps  = fp(kvs, d*1000, d1*1000, d2*1000)) %>% 
    mutate( flps = flp(kvs, fls, d*1000, d1*1000, d2*1000)) %>% 
    select( typ, kv.b, kv.d, kv.e, kvs, zvs,  fls, fps, flps)
  
  # Calculation characteristics
  
  base_data <- base_data %>% 
    mutate( dp = (p1 - p2),
            kv = kv(p1, p2, flow, temp)) %>% 
    mutate( zeta  = zeta_vaule(d, kv),
            sig_1 = sigma_1(p1, p2, elev, temp)) 
  
  data_analyse <- cylinder %>% 
    mutate(data = list(base_data)) %>% 
    unnest(data) %>% 
    mutate(kv_kvs = ifelse(kv > kvs, NA, kv/kvs), position = 0)
  
  for(i in c(1:length(data_analyse$kv_kvs))){
    if(is.na(data_analyse$kv_kvs[i])){
      data_analyse$position[i] <- NA
    } else {
      data_analyse$position[i] <- inv_LL3(data_analyse$kv_kvs[i], 
                                          data_analyse$kv.b[i],
                                          data_analyse$kv.d[i],
                                          data_analyse$kv.e[i])
    }
  }
  
  data_analyse <- data_analyse %>% 
    mutate( flp    = ifelse(kv > kvs, NA, fl_function( position, kv.b, kv.d, kv.e, flps)),
            Sig_i  = ifelse(kv > kvs, NA, Sigma_i( position,  kv.b,  kv.d,  kv.e, flps)),  # Incipient Cavitation
            Sig_c  = ifelse(kv > kvs, NA, Sigma_c( position,  kv.b,  kv.d,  kv.e, flps)),  # Constant Cavitation
            Sig_mv = ifelse(kv > kvs, NA, Sigma_mv( position,  kv.b,  kv.d,  kv.e, flps)), #  Maximum Vibration Cavitation
            regime = ifelse(kv > kvs, NA, cavtation_regime(position,  kv.b,  kv.d,  kv.e, flps, sig_1)))
  

  data_analyse <- data_analyse %>%
    group_by(typ, kv.b, kv.d, kv.e, zvs, kvs, fls, fps, flps) %>% 
    nest()

  
```

  1. The flow should be a constant `r flow` $m^3/h$ (`r flow/3600` $m^3/s$).
  2. The control valve has a diameter of `r round(d*1000,0)` mm and is smaller than the diameters of the pipe upstream (`r round(d1*1000,0)` mm) and downstream (`r round(d2*1000,0)` mm)  of the valve.

Service Conditions:

  - Fluid : Water 
  - Water Temperature: `r temp` °C
  - Vapour pressure: `r round(vapour_pressure(temp),3)` $bar$
  - Altitude: `r elev` msal.
  - Atm. Pressure: `r round(atm_pressure(masl = elev),3)` $bar$
  - Diameter Pipe Upstream $D_1$: `r round(d1*1000,0)` mm  
  - Diameter Pipe Downstream $D_2$: `r round(d2*1000,0)` mm  
  - Diameter Valve $d$: `r round(d*1000,0)` mm  
  - Valve: Plunger Valve
  
The maximum mean and minimum pressures calculated and the valve (upstream and downstream) for the necessary constant flow can be seen in table  \@ref(tab:pressures-Example-02). 


```{r pressures-Example-02, echo=FALSE}
  base_data %>% 
    select(measurement, p1, p2, flow, dp) %>% 
    kbl(caption ="Pressures and flow data at the valve",
        col.names = c("Measu.", "$P_1$" , "$P_2$", "$Flow$", "$\\Delta{P}$")) %>% 
    add_header_above(c("", "$bar$", "$bar$", "$m^3/h$", "$bar$")) %>% 
    kable_classic( bootstrap_options = "striped", 
                   full_width = F, position = "left")
```

The different types of Control Characteristic Parameters for the valve can be seen in the table \@ref(tab:valvetyp-Example-02) and in the figures \@ref(fig:kvkvs-typ-01-Example-2),  \@ref(fig:kvkvs-typ-02-Example-2), and  \@ref(fig:kvkvs-typ-03-Example-2).

```{r valvetyp-Example-02, echo=FALSE}
  cylinder %>%
    select(typ, kvs, zvs, fls) %>% 
    kbl(caption ="Parameter of the Control Characteristic",
        col.names = c("$Typ$", "$K_{v_s}$", "$\\zeta_{v_s}$", "$F_{L_s}$"),
        digits = c(0, 0, 2, 2)) %>% 
    kable_classic( bootstrap_options = "striped", full_width = F, position = "left") 
    
```


```{r kvkvs-typ-01-Example-2, echo=FALSE, fig.cap="Valve Characteristics valve typ 01"}

  # https://cran.r-project.org/web/packages/egg/vignettes/Ecosystem.html

  i <- 1
  p1 <- plot_kv(cylinder$kv.b[i], cylinder$kv.d[i], cylinder$kv.e[i], d, cylinder$zvs[i], cylinder$typ[i]) 
  p2 <- plot_zv(cylinder$kv.b[i], cylinder$kv.d[i], cylinder$kv.e[i], cylinder$zvs[i], cylinder$typ[i])
  grid.arrange(p1, p2, nrow = 1)
  
```

```{r kvkvs-typ-02-Example-2, echo=FALSE, fig.cap="Valve Characteristics valve typ 02"}

  i <- 2
  p1 <- plot_kv(cylinder$kv.b[i], cylinder$kv.d[i], cylinder$kv.e[i], d, cylinder$zvs[i], cylinder$typ[i]) 
  p2 <- plot_zv(cylinder$kv.b[i], cylinder$kv.d[i], cylinder$kv.e[i], cylinder$zvs[i], cylinder$typ[i])
  grid.arrange(p1, p2, nrow = 1)
  
```

```{r kvkvs-typ-03-Example-2, echo=FALSE, fig.cap="Valve Characteristics valve typ 03"}

  i <- 3
  p1 <- plot_kv(cylinder$kv.b[i], cylinder$kv.d[i], cylinder$kv.e[i], d, cylinder$zvs[i], cylinder$typ[i]) 
  p2 <- plot_zv(cylinder$kv.b[i], cylinder$kv.d[i], cylinder$kv.e[i], cylinder$zvs[i], cylinder$typ[i])
  grid.arrange(p1, p2, nrow = 1)
  
```


## Step 3. Determine the piping geometry factor ($F_{P_s}$) and the combined liquid pressure recovery factor ($F_{LP_s}$).

```{r valve-with-fittings2, echo=FALSE, fig.cap = 'Valve with fittings [@wagner2008]', fig.width = 10, fig.asp = 0.7, fig.align = "center"}
  file_image <- here::here("image", "valve_with_fittings.PNG")
  knitr::include_graphics(file_image)
```

according to equation \@ref(eq:Fp) :

$$
  F_{P} = \left( \frac{\sum{\zeta} \cdot K_v^2}{0.0016 \cdot d^4} + 1 \right)^{-1/2}
$$
and equation \@ref(eq:Flp) :

$$
  F_{LP} = F_L \cdot \left(\frac{\sum{\zeta} \cdot F_L^2 \cdot K_v^2}{0.0016 \cdot d^4} + 1 \right)^{-1/2}
$$

Where the factor $\sum{\zeta}$ \@ref(eq:sumZ) is the algebraic sum of the effective velocity head coefficients of all fittings attached to but not including the valve. For instance,

$$
  \sum{\zeta} = (\zeta_{1} + \zeta_{2}) + (\zeta_{B_1} - \zeta_{B_2})
$$

where

Resistance coefficient of the valve in the inlet \@ref(eq:Z1): 
    $\zeta_{1} = 0.5 \cdot \left( 1 - \left( \frac{d}{D_1}\right)^2 \right)^2 =$ `r 0.5*(1-(d/d1)^2)^2`
  
Resistance coefficient of the valve in the outlet \@ref(eq:Z2): 
    $\zeta_{2} = \left( 1 - \left( \frac{d}{D_2}\right)^2 \right)^2 =$ `r (1-(d/d2)^2)^2`
  
Bernoulli pressure number in the valve inlet \@ref(eq:Zb1): 
    $\zeta_{B_1} = 1- \left( \frac{d}{D_1}  \right)^4 =$  `r 1-(d/d1)^4`
  
Bernoulli pressure number in the valve outlet \@ref(eq:Zb2): 
    $\zeta_{B_2} = 1- \left( \frac{d}{D_2}  \right)^4 =$ `r 1-(d/d2)^4`

$\sum{\zeta} =$ `r resistance_coefficient(d, d1, d2)`


so:

```{r flp-Example-02, echo=FALSE}
  data_analyse %>% ungroup() %>% 
    select(typ, kvs, zvs, fls, fps, flps) %>% 
    kbl(caption ="Piping geometry factor and the combined liquid pressure recovery factor",
        col.names = c("$Typ$", "$K_{v_s}$", "$\\zeta_{v_s}$", "$F_{L_s}$", "$F_{P_s}$", "$F_{LP_s}$"),
        digits = c(0, 0, 2, 2, 2, 2)) %>% 
    kable_classic( bootstrap_options = "striped", full_width = F, position = "left") 
    
```

## Step 4. Calculation of needes Kv, Zv, sigma, valve position, and  kv/kvs

### With the valve  `r data_analyse$typ[1]`

```{r positions1-Example-02, echo=FALSE}
  data_analyse$data[[1]] %>%  
    mutate( flow = scales::comma(flow),
            kv = scales::comma(kv),
            sig_1 = round(sig_1,2),
            kv_kvs   = scales::percent(kv_kvs), 
            position = scales::percent(position/100)) %>% 
    select(measurement, p1, p2, flow, dp, kv, sig_1, kv_kvs, position) %>% 
    kbl(caption ="Pressures and flow data at the valve",
        col.names = c("Measu.", "$P_1$" , "$P_2$", "$Flow$", "$\\Delta{P}$", "$K_v$", "$\\sigma_{1}$", "$K_v/K_{vs}$", "Opening")) %>% 
    add_header_above(c("", "$bar$", "$bar$", "$m^3/h$", "$bar$", "$m^3/h$", "--", "%", "%")) %>% 
    kable_classic( bootstrap_options = "striped", 
                   full_width = F, position = "left") %>%
    footnote( symbol = c("NA values in $K_v/K_{vs}$, this means that the required flow capacity is not available.") )
```



```{r kv-kvs1-Example-2, echo=FALSE, fig.cap="Inherent Characteristic of the valve", warning=TRUE}

  i <- 1
  
  data_points <- data_analyse$data[[i]] %>% 
    select( position, kv_kvs) %>% 
    mutate( kv_kvs = kv_kvs * 100) %>% 
    rename( x = position, y = kv_kvs )
  
  segment_data <- tibble ( x = data_points$x,
                           y    = data_points$y*0,
                           xend = data_points$x,
                           yend = data_points$y) %>% 
                  add_row( x    = data_points$x*0,
                           y    = data_points$y,
                           xend = data_points$x,
                           yend = data_points$y)
  
  
  plot_kv_kvs( data_analyse[i,]$kv.b, data_analyse[i,]$kv.d, 
               data_analyse[i,]$kv.e, data_analyse[i,]$typ) +
    geom_point( data = data_points, 
                aes( x = x, y = y ), 
                colour = "black", size = 2) +
    geom_segment(data = segment_data, aes(x = x, y = y, xend = xend, yend = yend), linetype = "dotted")

```


### With the valve  `r data_analyse$typ[2]`

```{r positions2-Example-02, echo=FALSE}

  data_analyse$data[[2]] %>%  
    mutate( flow = scales::comma(flow),
            kv = scales::comma(kv),
            sig_1 = round(sig_1,2),
            kv_kvs   = scales::percent(kv_kvs), 
            position = scales::percent(position/100)) %>% 
    select(measurement, p1, p2, flow, dp, kv, sig_1, kv_kvs, position) %>% 
    kbl(caption ="Pressures and flow data at the valve",
        col.names = c("Measu.", "$P_1$" , "$P_2$", "$Flow$", "$\\Delta{P}$", "$K_v$", "$\\sigma_{1}$", "$K_v/K_{vs}$", "Opening")) %>% 
    add_header_above(c("", "$bar$", "$bar$", "$m^3/h$", "$bar$", "$m^3/h$", "--", "%", "%")) %>% 
    kable_classic( bootstrap_options = "striped", 
                   full_width = F, position = "left")%>%
    footnote( symbol = c("NA values in $K_v/K_{vs}$, this means that the required flow capacity is not available.") )

```


```{r kv-kvs2-Example-2, echo=FALSE, fig.cap="Inherent Characteristic of the valve", warning=TRUE}

  i <- 2
  
  data_points <- data_analyse$data[[i]] %>% 
    select( position, kv_kvs) %>% 
    mutate( kv_kvs = kv_kvs * 100) %>% 
    rename( x = position, y = kv_kvs )
  
  segment_data <- tibble ( x = data_points$x,
                           y    = data_points$y*0,
                           xend = data_points$x,
                           yend = data_points$y) %>% 
                  add_row( x    = data_points$x*0,
                           y    = data_points$y,
                           xend = data_points$x,
                           yend = data_points$y)
  
  
  plot_kv_kvs( data_analyse[i,]$kv.b, data_analyse[i,]$kv.d, 
               data_analyse[i,]$kv.e, data_analyse[i,]$typ) +
    geom_point( data = data_points, 
                aes( x = x, y = y ), 
                colour = "black", size = 2) +
    geom_segment(data = segment_data, aes(x = x, y = y, xend = xend, yend = yend), linetype = "dotted")

```



### With the valve  `r data_analyse$typ[3]`

```{r positions3-Example-02, echo=FALSE}

  data_analyse$data[[3]] %>%  
    mutate( flow = scales::comma(flow),
            kv = scales::comma(kv),
            sig_1 = round(sig_1,2),
            kv_kvs   = scales::percent(kv_kvs), 
            position = scales::percent(position/100)) %>% 
    select(measurement, p1, p2, flow, dp, kv, sig_1, kv_kvs, position) %>% 
    kbl(caption ="Pressures and flow data at the valve",
        col.names = c("Measu.", "$P_1$" , "$P_2$", "$Flow$", "$\\Delta{P}$", "$K_v$", "$\\sigma_{1}$", "$K_v/K_{vs}$", "Opening")) %>% 
    add_header_above(c("", "$bar$", "$bar$", "$m^3/h$", "$bar$", "$m^3/h$", "--", "%", "%")) %>% 
    kable_classic( bootstrap_options = "striped", 
                   full_width = F, position = "left")%>%
    footnote( symbol = c("NA values in $K_v/K_{vs}$, this means that the required flow capacity is not available.") )

```

## Step 5. Calcultion of the Cavitation's borders.

### For the valve  `r data_analyse$typ[1]`

```{r sigma1-Example-02, echo=FALSE}
  data_analyse$data[[1]] %>%  
    mutate( sig_1  = round(sig_1,  2),
            Sig_i  = round(Sig_i,  2),
            Sig_c  = round(Sig_c,  2), 
            Sig_mv = round(Sig_mv, 2),
            flp    = round(flp,  2),
            position = scales::percent(position/100)) %>% 
    select(measurement, p1, p2, dp, position, flp, sig_1, Sig_i, Sig_c, Sig_mv, regime) %>% 
    kbl(caption ="Sigma value and cavitation's borders",
        col.names = c("Measu.", "$P_1$" , "$P_2$", "$\\Delta{P}$",
                      "Opening", "$F_{LP}$", "$\\sigma_{1}$", "$\\sigma_{i}$", "$\\sigma_{c}$", "$\\sigma_{mv}$", "Regime")) %>% 
    add_header_above(c("", "$bar$", "$bar$", "$bar$", "%", "", "", "", "", "", "")) %>% 
    kable_classic( bootstrap_options = "striped",
                   full_width = F, position = "left") %>%
    footnote( symbol = c("NA values in $position$ and $\\sigma$, this means that the required flow capacity is not available.") )
```

```{r sigma1-Example-2, echo=FALSE, warning=FALSE, fig.cap="Sigma"}

  i <- 1

  data_points <- data_analyse$data[[i]] %>% 
    select( position, Sig_i) %>% 
    rename( x = position, y = Sig_i )
  
  segment_data <- tibble ( x = data_points$x,
                           y    = data_points$y*0,
                           xend = data_points$x,
                           yend = data_points$y) %>% 
                  add_row( x    = data_points$x*0,
                           y    = data_points$y,
                           xend = data_points$x,
                           yend = data_points$y)
    
  plot_sigma( data_analyse[i,]$kv.b, data_analyse[i,]$kv.d, 
              data_analyse[i,]$kv.e, data_analyse[i,]$fls, 
              data_analyse[i,]$typ) +
    geom_point( data = data_points, aes( y = y, x = x ), colour = "black", size = 2) +
    geom_segment(data = segment_data, aes(x = x, y = y, xend = xend, yend = yend), linetype = "dotted")  

```


### For the valve  `r data_analyse$typ[2]`

```{r sigma2-Example-02, echo=FALSE}
  data_analyse$data[[2]] %>%  
    mutate( sig_1  = round(sig_1,  2),
            Sig_i  = round(Sig_i,  2),
            Sig_c  = round(Sig_c,  2), 
            Sig_mv = round(Sig_mv, 2),
            flp    = round(flp,  2),
            position = scales::percent(position/100)) %>% 
    select(measurement, p1, p2, dp, position, flp, sig_1, Sig_i, Sig_c, Sig_mv, regime) %>% 
    kbl(caption ="Sigma value and cavitation's borders",
        col.names = c("Measu.", "$P_1$" , "$P_2$", "$\\Delta{P}$",
                      "Opening", "$F_{LP}$", "$\\sigma_{1}$", "$\\sigma_{i}$", "$\\sigma_{c}$", "$\\sigma_{mv}$", "Regime")) %>% 
    add_header_above(c("", "$bar$", "$bar$", "$bar$", "%", "", "", "", "", "", "")) %>% 
    kable_classic( bootstrap_options = "striped",
                   full_width = F, position = "left") %>%
    footnote( symbol = c("NA values in $position$ and $\\sigma$, this means that the required flow capacity is not available.") )
```


```{r sigma2-Example-2, echo=FALSE, warning=FALSE, fig.cap="Sigma"}

  i <- 2

  data_points <- data_analyse$data[[i]] %>% 
    select( position, Sig_i) %>% 
    rename( x = position, y = Sig_i )

    segment_data <- tibble ( x = data_points$x,
                           y    = data_points$y*0,
                           xend = data_points$x,
                           yend = data_points$y) %>% 
                  add_row( x    = data_points$x*0,
                           y    = data_points$y,
                           xend = data_points$x,
                           yend = data_points$y)

  plot_sigma( data_analyse[i,]$kv.b, data_analyse[i,]$kv.d, 
              data_analyse[i,]$kv.e, data_analyse[i,]$fls, 
              data_analyse[i,]$typ) +
    geom_point( data = data_points, aes( y = y, x = x ), colour = "black", size = 2)  +
    geom_segment(data = segment_data, aes(x = x, y = y, xend = xend, yend = yend), linetype = "dotted")   

```


### For the valve  `r data_analyse$typ[3]`

```{r sigma3-Example-02, echo=FALSE}
  data_analyse$data[[3]] %>%  
    mutate( sig_1  = round(sig_1, 2),
            Sig_i  = round(Sig_i, 2),
            Sig_c  = round(Sig_c, 2), 
            Sig_mv = round(Sig_mv, 2),
            flp    = round(flp, 2),
            position = scales::percent(position/100)) %>% 
    select(measurement, p1, p2, dp, position, flp, sig_1, Sig_i, Sig_c, Sig_mv, regime) %>% 
    kbl(caption ="Sigma value and cavitation's borders",
        col.names = c("Measu.", "$P_1$" , "$P_2$", "$\\Delta{P}$",
                      "Opening", "$F_{LP}$", "$\\sigma_{1}$", "$\\sigma_{i}$", "$\\sigma_{c}$", "$\\sigma_{mv}$", "Regime")) %>% 
    add_header_above(c("", "$bar$", "$bar$", "$bar$", "%", "", "", "", "", "", "")) %>% 
    kable_classic( bootstrap_options = "striped",
                   full_width = F, position = "left") %>%
    footnote( symbol = c("NA values in $position$ and $\\sigma$, this means that the required flow capacity is not available.") )
```

## Step 6. Conclusion

the initial conclusions would be the following: 

  - Only type 1 and 2 valves meet the required conditions; 
  - With the information of the minimum, average and maximum values, only a first approximation of the behavior of the valve is achieved. And it is used for the preselection of the type of valve required. Depending on the importance of the valve for the operation of the system, it is recommended to know all the values corresponding to all the operating scenarios. 
  - This exercise is derived from the actual selection of two control valves on a water treatment plant. The valves are used to control the backwashing of the filters, and for this reason, it is extremely important to review all operating scenarios (see Example 3).
  
